---
title             : "Into the UFC: Best Fighter, Striker, Grappler, and Entertainer"
shorttitle        : "UFC Stats"

author: 
  - name          : "Stanley Go"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "smg421@scarletmail.rutgers.edu"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Rutgers University"

authornote: |
  This project was submitted on `r Sys.Date()`

abstract: 
  
  This project delves into an in-depth analysis of UFC fighter statistics to identify and recognize excellence in various performance categories. The primary objective is to pinpoint the best fighters in distinct areas such as striking, grappling, knockout ability, and overall entertainment value. The study further aims to determine the greatest of all time (GOAT) by evaluating fighters based on their win-loss ratios. Leveraging a comprehensive dataset encompassing fighter metrics, the analysis employs key indicators such as significant strikes rate, takedown success rate, and total knockdowns. The anticipated results include detailed insights into the best striker, grappler, KOer, and entertainer, contributing to a comprehensive understanding of individual strengths within the competitive realm of UFC. Additionally, the study seeks to establish the GOAT by considering the historical performance of fighters with a minimum threshold of 10 matches. Through meticulous analysis and visualization, this project aims to offer a nuanced perspective on the unparalleled skills and accomplishments of UFC fighters across diverse categories.
  


  
keywords          : "Best, GOAT, Fighter, Performance, UFC"
wordcount         : "160"

bibliography      : "r-references.bib"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
library("tidyverse")
r_refs("r-references.bib")
```

```{r load-data, echo=FALSE, results='hide'}
load("C:/Users/stanl/OneDrive/Documents/GitHub/412/FinalProject/ufc_stats.rda")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction: 
  
  - Background:
      - Brief overview of the UFC and its significance in the world of mixed martial arts (MMA).
      - Growing interest in understanding and analyzing fighter performance.

  - Research Focus:
      - Exploration of UFC fighter statistics to recognize excellence in various performance categories.

  - Big Question:
      - The Ultimate Fighting Championship (UFC) hosts a wide range of fighters with diverse skill sets, including striking, grappling, and entertainment value. The objective of this project is to systematically analyze UFC fighter data to determine who excels in each category, identify the best overall fighters, and predict fight outcomes and potential earnings. We will aim our analysis to provide insights into fighter performance, audience appeal, and financial success within the UFC.

  - Objectives:
      - Identification of the best fighters in specific areas:
          - Best Striker
          - Best Grappler
          - Best KOer
          - Best Entertainer

  - Key Questions:
      - Research questions guiding the analysis:
          - Who are the best strikers, grapplers, and KOers in UFC?
          - What factors contribute to a fighter's entertainment value?
          - Who is considered the greatest of all time in UFC?

  - Methodology:
      - Planned analysis includes calculating key metrics such as significant strikes rate, takedown success rate, and total knockdowns.

  - Anticipated Results:
      - Expected outcomes involve insights into the best performers in each category, contributing to a nuanced understanding of individual strengths.

  - Significance:
      - Emphasis on the significance of recognizing and appreciating excellence within the competitive realm of UFC.

  - Overview of the Document:
      - Brief mention of the subsequent sections, including the methodology, results, and discussion.


# Methods
## Data Collection

- **Data Source:**
  - Utilize the UFC Stats dataset, containing comprehensive fighter metrics, including but not limited to knockdowns, significant strikes, takedowns, and fight outcomes.

## Analysis Approach

1. **Identifying Best Striker:**
   - Calculate the significant strikes rate for each fighter:
     \[ \text{Significant Strikes Rate} = \frac{\text{Total Significant Strikes Landed}}{\text{Total Significant Strikes Attempted}} \]
   - Determine the fighter with the highest significant strikes rate as the best striker.

2. **Identifying Best Grappler:**
   - Compute the takedown success rate for each fighter:
     \[ \text{Takedown Success Rate} = \frac{\text{Total Successful Takedowns}}{\text{Total Takedown Attempts}} \]
   - Identify the fighter with the highest takedown success rate as the best grappler.

3. **Identifying Best KOer:**
   - Sum the total knockdowns for each fighter.
   - Determine the fighter with the highest total knockdowns as the best KOer.

4. **Identifying Best Entertainer:**
   - Develop a composite metric considering both significant strikes and successful takedowns.
   - Determine the fighter with the highest composite metric as the best entertainer.

5. **Determining Greatest of All Time (GOAT):**
   - Filter fighters with a minimum threshold of 10 matches.
   - Calculate the win-loss ratio for each fighter:
     \[ \text{Win-Loss Ratio} = \frac{\text{Total Wins}}{\text{Total Wins + Total Losses}} \]
   - Identify the fighter with the highest win-loss ratio as the GOAT.
   - Didn't based it on striking + grappling efficiency as it is disingenuous to a better      metric which is W/L ratio.  

## Visualization

- **Figure Creation:**
  - Develop bar plots for each category (Best Striker, Best Grappler, Best KOer, Best Entertainer, and GOAT) to visually represent the analysis results.

## Limitations

- **Limitations of the Analysis:**
  - Acknowledge potential limitations such as data accuracy, sample size variations, and the subjectivity of composite metrics.

## Participants
- UFC Fighters

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses. And, ufc_stats.rda.


# Results
## Planned Analysis

Our analysis focused on several key aspects of UFC fighter performance, including:

- **Best Striker:** We calculated the total significant strikes landed for each fighter and identified the top performers.

- **Best Grappler:** Utilizing the grappling efficiency metric, we determined the fighters with the highest success rate in takedowns.

- **Best KOer:** We considered knockdowns, significant strikes, and other metrics to identify fighters with exceptional knockout abilities.

- **Best Entertainer:** A composite metric, combining significant strikes, takedowns, and knockdowns, helped us identify fighters who entertain the audience.

- **Best PPV Seller:** We calculated the average PPV revenue per fight for each fighter, focusing on financial success.

- **Greatest of All Time (GOAT):** Based on win-loss ratio and total wins, we identified the top fighters with at least 15 fights. 


# Discussion
The analysis of UFC fighter statistics has provided valuable insights into various performance categories, shedding light on the strengths and capabilities of individual fighters within the competitive realm of mixed martial arts (MMA). This discussion will systematically address the research questions posed in the introduction, connecting the findings to the broader context of UFC fighter performance.

## Best Striker in UFC
The identification of the best striker was based on the significant strikes rate, calculated as the ratio of total significant strikes landed to attempted. The analysis revealed [Name] as the fighter with the highest striking efficiency, showcasing exceptional skill in landing significant strikes. This result aligns with the goal of recognizing technical proficiency in striking, providing fans and analysts with a nuanced perspective on the stand-up game.

## Best Grappler in UFC
Grappling efficiency, defined as the ratio of successful takedowns to attempted takedowns, was utilized to pinpoint the best grappler. [Name] emerged as the fighter with the highest grappling efficiency, showcasing dominance in ground control and takedown success. This insight contributes to the understanding of diverse skill sets among UFC fighters, emphasizing the importance of both striking and grappling in MMA.

## Greatest of All Time (GOAT) Analysis
The determination of the Greatest of All Time (GOAT) involved assessing fighters based on their win-loss ratio and total wins, with a minimum fight threshold. The analysis identified [Name] as the top-ranked fighter, emphasizing sustained success and overall excellence. The scatter plot visually represents the relationship between total wins and the win-loss ratio for the top 20 GOAT fighters. This graphical representation adds depth to the analysis, allowing for a quick comparison of fighters' career trajectories.

## Limitations and Future Directions
Despite the insightful findings, this analysis has its limitations. The reliance on statistical metrics may overlook intangible factors influencing a fighter's performance. Additionally, variations in data accuracy and the subjective nature of composite metrics pose challenges. Future research could delve into incorporating qualitative data and exploring advanced statistical models to offer a more comprehensive understanding of UFC fighter excellence.

In conclusion, the discussion has provided a comprehensive interpretation of the data, addressing the research questions and highlighting the strengths and limitations of the analysis. The nuanced insights into striking, grappling, and overall fighter performance contribute to the ongoing discourse surrounding the diverse skill sets within the UFC. The discussion sets the stage for future research, encouraging a more holistic approach to assessing fighter excellence in MMA.


## Planned Figure and Table
 *Figure: Top 20 GOAT (Win-Loss Ratio and Total Wins)*

```{r}
ufc_stats_filtered <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(
    total_wins = sum(winner == "W"),
    total_losses = sum(winner == "L"),
    total_fights = n_distinct(id)
  ) %>%
  filter(total_fights >= 15)  

ufc_stats_filtered <- ufc_stats_filtered %>%
  mutate(win_loss_ratio = total_wins / (total_wins + total_losses))

# Identify Top 20 GOAT
top_20_goat_filtered <- ufc_stats_filtered %>%
  arrange(desc(win_loss_ratio), desc(total_wins)) %>%
  head(20)

# Plot the Top 20 GOAT as a scatter plot with names
ggplot(top_20_goat_filtered, aes(x = total_wins, y = win_loss_ratio, label = fighter)) +
  geom_point(color = "blue") +  
  geom_text(vjust = -0.5, hjust = 0.5) +  
  labs(title = "Top 20 GOAT (Based on Win-Loss Ratio and Total Wins)", x = "Total Wins", y = "Win-Loss Ratio") +
  theme_minimal()
```

  - In this figure, we compare fighters based on their win-loss record. We combine total     wins of fighters and divide that by their total record which includes their wins +       losses to calculate their ratio. This tells us who has the best percentage of success
    in the octagon.

 *Figure: Top 20 Entertainers* 
```{r}
# Identifying Best Entertainers
best_entertainer_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(entertainment_metric = sum(significant_strikes_landed) + sum(takedown_successful)) %>%
  arrange(desc(entertainment_metric)) %>%
  head(20)  # Adjust the number to 20 for the top 20 entertainers

# Plot the Best Entertainer (Top 20)
ggplot(best_entertainer_data, aes(x = reorder(fighter, -entertainment_metric), y = entertainment_metric)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Best Entertainer (Composite Metric) - Top 20", x = "Fighter", y = "Composite Metric") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

  - In this figure, we calculate the entertainment metric based on the strikes landed and    takedowns landed. This is a good metric as viewers tend to be drawn towards              significant hits on the head or body. We then used that and apply it to 20 UFC           fighters

 *Figure: Top 20 PPV Seller* 
```{r}
# Identify Overall Best PPV Seller
best_ppv_seller_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(total_ppv_revenue = sum(attendance * 59.99)) %>%
  arrange(desc(total_ppv_revenue)) %>%
  head(20)  # Adjust the number to 20 for the top 20 PPV sellers

# Plot the Best PPV Sellers (Top 20)
ggplot(best_ppv_seller_data, aes(x = reorder(fighter, -total_ppv_revenue), y = total_ppv_revenue)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Best PPV Seller (Total PPV Revenue) - Top 20", x = "Fighter", y = "Total PPV Revenue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
 - In this figure, we calculate who the best PPV seller is in the UFC. We mainly use        attendance as the key metric and multiply that by 59.99 to get the total revenue made    by the UFC based on that fighter.
 
 *Figure: Top 20 Koer* 
```{r}
# Identifying Best KOers
best_koer_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(total_knockdowns = sum(knockdowns)) %>%
  arrange(desc(total_knockdowns)) %>%
  head(20)  # Adjust the number to 20 for the top 20 KOers

# Plot the Best KOer (Top 20)
ggplot(best_koer_data, aes(x = reorder(fighter, -total_knockdowns), y = total_knockdowns)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Best KOer (Total Knockdowns) - Top 20", x = "Fighter", y = "Total Knockdowns") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

  - In this figure, we calculate the most accurate fighter based on knockdowns. Knockdown     is a good indicator of how accurate and strong a fighter is. We use sum of variable      knockdowns to see how much knockdowns a fighter has in their career. 

 *Figure: Top 20 Striker*
```{r}
# Identifying Best Strikers
best_striker_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(significant_strikes_rate = sum(significant_strikes_landed) / sum(significant_strikes_attempted)) %>%
  arrange(desc(significant_strikes_rate)) %>%
  head(20)  # Adjust the number to 20 for the top 20 strikers

# Plot the Best Striker (Top 20)
ggplot(best_striker_data, aes(x = reorder(fighter, -significant_strikes_rate), y = significant_strikes_rate)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Best Striker (Significant Strikes Rate) - Top 20", x = "Fighter", y = "Significant Strikes Rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

  -   In this figure, we calculate the best striker based on the significance strikes          landed rate. To do this, we sum the amount landed and strikes attempted to find the       rate. It is a good measurement since it finds out who the most accurate and              effective striker is. 

 *Figure: Top 20 Grapplers*
```{r}
# Identifying Best Grapplers
best_grappler_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(grappling_efficiency = sum(takedown_successful) / sum(takedown_attempted)) %>%
  arrange(desc(grappling_efficiency)) %>%
  head(20)  # Adjust the number to 20 for the top 20 grapplers

# Plot the Best Grappler (Top 20)
ggplot(best_grappler_data, aes(x = reorder(fighter, -grappling_efficiency), y = grappling_efficiency)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(title = "Best Grappler (Grappling Efficiency) - Top 20", x = "Fighter", y = "Grappling Efficiency") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

  - In this figure, we find the best grappler by using grappling efficiency. We calculate     using the sum of takedown success over the sum of takedown attempted.
 *Table KOer*
```{r create-table, results='asis'}

# Identifying Best KOers
best_koer_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(total_knockdowns = sum(knockdowns)) %>%
  arrange(desc(total_knockdowns)) %>%
  head(20)  # Adjust the number to 20 for the top 20 KOers

# Create a table
best_koer_table <- best_koer_data %>%
  knitr::kable(digits = 0, align = "c")

best_koer_table
```

 *Table Striker* 

```{r create-striking-efficiency-table, results='asis'}

# Identifying Top Fighters based on Striking Efficiency
best_striking_efficiency_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(striking_efficiency = sum(significant_strikes_landed) / sum(significant_strikes_attempted)) %>%
  arrange(desc(striking_efficiency)) %>%
  head()  # Adjust the number based on your preferences


best_striking_efficiency_table <- best_striking_efficiency_data %>%
  knitr::kable(digits = 2, align = "c")

best_striking_efficiency_table

```

 *Table for Grappler*
```{r create-grappling-efficiency-table, results='asis'}

# Identifying Top Fighters based on Grappling Efficiency
best_grappling_efficiency_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(grappling_efficiency = sum(takedown_successful) / sum(takedown_attempted)) %>%
  arrange(desc(grappling_efficiency)) %>%
  head()  # Adjust the number based on your preferences


best_grappling_efficiency_table <- best_grappling_efficiency_data %>%
  knitr::kable(digits = 2, align = "c")

best_grappling_efficiency_table

```

 *Table for PPV Seller*
 
```{r create-ppv-seller-table, results='asis'}


# Identify Overall Best PPV Seller
best_ppv_seller_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(total_ppv_revenue = sum(attendance * 59.99)) %>%
  arrange(desc(total_ppv_revenue)) %>%
  head(20)  # Adjust the number based on your preferences


best_ppv_seller_table <- best_ppv_seller_data %>%
  knitr::kable(digits = 2, align = "c")

best_ppv_seller_table

```

 *Table for Entertainer*
```{r create-entertainer-table, results='asis'}


# Identify Overall Best Entertainer
best_entertainer_data <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(entertainment_metric = sum(significant_strikes_landed) + sum(takedown_successful)) %>%
  arrange(desc(entertainment_metric)) %>%
  head(20)  # Adjust the number based on your preferences


best_entertainer_table <- best_entertainer_data %>%
  knitr::kable(digits = 2, align = "c")

best_entertainer_table


```

 *Table for the GOAT*
```{r create-goat-table, results='asis'}

# Filter and Calculate GOAT Metrics
ufc_stats_filtered <- ufc_stats %>%
  group_by(fighter) %>%
  summarise(
    total_wins = sum(winner == "W"),
    total_losses = sum(winner == "L"),
    total_fights = n_distinct(id)
  ) %>%
  filter(total_fights >= 15)  # Adjust the minimum number of fights as needed

ufc_stats_filtered <- ufc_stats_filtered %>%
  mutate(win_loss_ratio = total_wins / (total_wins + total_losses))

# Identify Top 20 GOAT
top_20_goat_filtered <- ufc_stats_filtered %>%
  arrange(desc(win_loss_ratio), desc(total_wins)) %>%
  head(20)


goat_table <- top_20_goat_filtered %>%
  knitr::kable(digits = 2, align = "c")

goat_table


```


# Discussion


\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
